\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\newtheorem{result}{Result}

\title{Pass-Through, Welfare, and Incidence under Product Returns\\
  in Perfect Competition}
\author{Draft notes}
\date{\today}

\begin{document}
\maketitle

%=============================================================================
\section{Introduction}
%=============================================================================

About one in five products sold online is returned.  Pass-through, tax
incidence, and the efficiency cost of taxation are well understood when
purchase settles who keeps the good.  In that classical setting, the demand
and supply curves are sufficient for consumer surplus, producer surplus, and
deadweight loss.  With returns, allocation is decided in two stages.  Buyers
first choose whether to purchase, then whether to keep or return after
receiving the item.  Never buying and buying then returning leave the same
final allocation---the consumer does not keep the good---yet their welfare
costs differ.  A returned purchase incurs shipping, hassle, and handling that
a non-purchase avoids, so demand and supply alone omit an economically
relevant margin.

What objects then replace the classical demand and supply curves, and how
misleading is an analysis that ignores returns?  We study these questions under
perfect competition with product returns and seller handling costs.  We
characterize pass-through, incidence, and welfare losses with returns, compare
them to the no-return benchmark, and identify where omitting the return margin
changes conclusions.

Consumer surplus depends only on two aggregates, gross demand and the return
rate.  Heterogeneity in tastes, match values, and return costs enters welfare
solely through these aggregates.  Producer surplus retains the usual
geometry---area to the left of the supply curve---but is evaluated at net
revenue, list price minus expected refunds and handling, rather than at list
price.

Because consumers pay the list price while firms respond to net revenue,
pass-through from cost shocks into list prices is governed by how net revenue
moves with the list price.  Tax neutrality fails for a different reason: the
keep-or-return decision depends on the refund price, which differs according
to which party remits the tax.  An analyst who ignores these two differences
and applies the classical formulas therefore misstates pass-through,
incidence, and deadweight loss whenever the return margin is active.  We show that pass-through of a one-dollar production-cost shock equals that
of a seller-remitted unit tax, but differs under buyer remittance.  The ratio
of handling to production-cost pass-through is governed only by the aggregate
return rate.  We also give an easy check, from the return-rate elasticity and
the ratio of handling to list price, for when the classical no-return
pass-through formula over- or underestimates the correct returns-adjusted
pass-through.

%=============================================================================
\section{Setting}
%=============================================================================

We begin with the market primitives: purchase and return behavior on the
consumer side, and competitive supply with return handling on the firm side.
A homogeneous product is sold at list price $p$.  Consumers indexed by $i$
differ in return hassle $\kappa_i$, the disutility from returning the product,
and in match distribution $G_i$ over post-purchase match quality $u$.  After
purchase, type $i$ draws $u\sim G_i$ and keeps the item whenever
$u-p\geq -\kappa_i$, earning surplus $u-p$, and returns it otherwise, earning
$-\kappa_i$.

Buying at price $p$ yields expected surplus
\[
  s_i(p)=\mathbb{E}_{u\sim G_i}\bigl[\max(u-p,\,-\kappa_i)\bigr].
\]
On the purchase margin, all consumer-side heterogeneity is summarized by
willingness to pay $v_i$, defined by $s_i(v_i)=0$, so type $i$ buys if and
only if $p\leq v_i$.  Across types, willingness to pay has distribution
$F(v)=\Pr(v_i\leq v)$ with density $f$, so gross demand is $D(p)=1-F(p)$.
Of the units sold at price $p$, a share $r(p)$ are returned.

To understand how $r(p)$ changes with price, write type $i$'s return
probability as $\rho_i(p)=\Pr(u<p-\kappa_i)$.  Willingness to pay is
sufficient for purchase but not for returns: types with the same $v_i$ may
still differ in $(\kappa_i,G_i)$.  Averaging within WTP cells,
\[
  \rho(v,p)\equiv\mathbb{E}\bigl[\rho_i(p)\bigm|v_i=v\bigr],
\]
so the aggregate return rate is the truncated mean
\[
  r(p)=\mathbb{E}\bigl[\rho(v_i,p)\bigm|v_i\geq p\bigr].
\]
Differentiating yields
\[
  r'(p)
  =\mathbb{E}\bigl[\partial\rho(v_i,p)/\partial p\bigm|v_i\geq p\bigr]
  +\frac{f(p)}{D(p)}\bigl(r(p)-\rho(p,p)\bigr),
\]
where $\rho(p,p)$ is the average return probability among types at the
purchase margin $v_i=p$.\footnote{With $\rho(v,p)$ as defined in the text,
$r(p)=D(p)^{-1}\int_p^\infty \rho(v,p)\,dF(v)$.  Differentiating and using
$D'(p)=-f(p)$ gives
\[
  r'(p)
  =D(p)^{-1}\int_p^\infty \partial\rho(v,p)/\partial p\,dF(v)
  +\frac{f(p)}{D(p)}\bigl(r(p)-\rho(p,p)\bigr),
\]
which is the display in the text.}  The first term is behavioral: a
higher keep threshold raises return probabilities among those who still buy.
The second is selection: a higher price drops cutoff types at the purchase
margin, and that term would be negative if those cutoff types return more than
the average purchaser.\footnote{Selection need not be negative for arbitrary
$(\kappa_i,G_i)$.  It is signed under an additive match specification
$u=v_i+\varepsilon$ with shock $\varepsilon$ of any distribution independent
of willingness to pay, and with return hassle likewise independent of $v$:
then $\rho(v,p)$ decreases in $v$, so $\rho(p,p)\geq r(p)$ and the selection
term is non-positive.}  Behavioral and selection effects can therefore pull
in opposite directions, and we do not impose a sign on $r'(p)$.

The product market is perfectly competitive.  Producer $j$ takes $p$ as given,
chooses output at cost $C_j(q_j)$, and incurs a net return handling cost $h$ on
each returned unit, equal to gross handling cost minus the scrap value of the
returned product.  This net cost may be negative, but we maintain $p+h>0$,
which holds whenever scrap value does not exceed list price.

%=============================================================================
\section{Consumer side}
%=============================================================================

We turn first to consumer surplus and show that only the aggregates $D$ and
$r$ enter.  Write $s(v,p)$ for expected surplus at list price $p$ among types
with willingness to pay $v$, so $s(v,v)=0$.  Consumer surplus integrates surplus
over all types who buy,
\begin{equation}
  CS(p)=\int_p^\infty s(v,p)\,dF(v).
\end{equation}
Differentiating in $p$ gives an extensive-margin term from types entering or
leaving the purchase margin and an intensive-margin term from lower surplus
among inframarginal buyers,
\[
  CS'(p)
  = -s(p,p)\,f(p)+\int_p^\infty s_p(v,p)\,dF(v),
\]
where $s_p(v,p)\equiv \partial s(v,p)/\partial p$.  On the extensive margin,
$s(p,p)=0$, so entry and exit contribute nothing at first order.  On the intensive margin, $s_p(v,p)$
equals minus the keep probability, because a higher price harms the consumer
only when she keeps the purchase,
\[
  s_p(v,p)=-\Pr(u\geq p-\kappa_i).
\]
Aggregating over buyers who purchase then gives
\[
  CS'(p)=\int_p^\infty s_p(v,p)\,dF(v)
  = -D(p)\bigl(1-r(p)\bigr),
\]
by the definitions of $D(p)$ and $r(p)$.  A small price change therefore satisfies
\begin{equation}
  dCS = -D(p)\bigl(1-r(p)\bigr)\,dp.
  \label{eq:dCS}
\end{equation}

The extensive- and intensive-margin decomposition makes the local formula
transparent.  Consumer surplus changes depend only on the aggregates $D(p)$ and
$r(p)$, both at each price and integrated along any path,
\[
  \Delta CS = -\int_{p_0}^{p_1} D(p)\bigl(1-r(p)\bigr)\,dp.
\]
Micro heterogeneity in price responses can be rich.  Price $p$ enters each
$s_i(p)$ nonlinearly and can shift participation, return decisions, and
inframarginal surplus differently across types.  Yet once $(D,r)$ are given, this
microstructure is irrelevant for $dCS$ and $\Delta CS$.  Two economies with
the same paths of $D(p)$ and $r(p)$ deliver the same consumer surplus loss,
whether high-$v_i$ types adjust return behavior more steeply than low-$v_i$
types or not.  Only kept units enter at first order, through $1-r(p)$, not
gross purchases $D(p)$ alone.

Let $\bar{p}$ denote the highest willingness to pay in the population, so
$D(\bar{p})=0$ and $CS(\bar{p})=0$.  Integrating the local formula then
rewrites consumer surplus as area under kept volume,
\begin{equation}
  CS(p)=\int_p^{\bar{p}} D(z)\bigl(1-r(z)\bigr)\,dz.
\end{equation}

%=============================================================================
\section{Supply side}
%=============================================================================

On the firm side, returns make list price an imperfect summary of producer
payoffs.  A gross sale brings revenue $p$ if the consumer keeps the item and
costs refund $p$ plus net handling $h$ if the item is returned.  Among gross
sales at list price $p$, share $r(p)$ are returned, so expected revenue per
gross unit---the \emph{effective marginal revenue}---is
\begin{equation}
  \mu(p,h) \equiv p - r(p)(p+h).
  \label{eq:mu-def}
\end{equation}
Under perfect competition, firm $j$ takes $p$ as given and therefore faces a
fixed number $\mu$ in its supply decision.  At equilibrium, $\mu=\mu(p,h)$.

Given $\mu$, firm $j$ chooses output to solve
\[
  \max_{q_j \geq 0}\;\; \mu\,q_j - C_j(q_j),
  \qquad
  \text{FOC: } C_j'(q_j)=\mu \;\;(\text{if } q_j>0),
\]
with $C_j(0)=0$.  Write $q_j(\mu)$ for firm $j$'s supply, set to zero when
$\mu$ is below its minimum marginal cost.  We assume each $C_j$ has strictly
increasing marginal cost, so $q_j'(\mu)>0$ for active firms.

Summing across firms defines market supply and industry profit,
\begin{equation}
  S(\mu) \equiv \sum_j q_j(\mu),
  \qquad
  PS(\mu) \equiv \sum_j \bigl[\mu\,q_j(\mu)-C_j\bigl(q_j(\mu)\bigr)\bigr].
  \label{eq:agg-def}
\end{equation}
Under that assumption, $S'(\mu)>0$.  Market supply therefore rises with
effective revenue, but not necessarily with list price.  Firms care about
$\mu$, and $\mu(p,h)$ can fail to rise with $p$ when refunds and handling
eat into the list-price gain.  Which firms are active at each $\mu$
determines both the market quantity and total profit.

Differentiating $PS(\mu)$ with respect to $\mu$ gives, for each firm,
\[
  \frac{d}{d\mu}\Bigl[\mu\,q_j-C_j(q_j)\Bigr]
  = q_j + q_j'(\mu)\,\bigl[\mu-C_j'(q_j)\bigr].
\]
Active firms satisfy $C_j'(q_j)=\mu$ at their chosen output, so the adjustment
term $q_j'(\mu)\,[\mu-C_j'(q_j)]$ is zero for each firm---equivalently, the
envelope theorem applied to each firm's profit problem gives
$d\pi_j/d\mu=q_j(\mu)$.  Summing over firms,
\[
  PS'(\mu)=\sum_j q_j(\mu)=S(\mu).
\]
With no production at $\mu=0$, it follows that industry profit equals the area
to the left of the market supply curve,
\begin{equation}
  PS(\mu)=\int_0^\mu S(z)\,dz.
  \label{eq:PS-integral}
\end{equation}
At equilibrium $\mu=\mu(p,h)$,
$PS=\int_0^{\mu(p,h)}S(z)\,dz$.

%=============================================================================
\section{Failure of tax neutrality}
%=============================================================================
\label{sec:tax-neutrality}

With firms responding to $\mu$ rather than list price, a natural question is
whether remittance of a unit tax remains neutral.  Weyl and Fabinger (2013)
show that without returns, buyer and seller remittance deliver the same
equilibrium.  With returns, we ask whether that equivalence survives by writing
the two clearing conditions at a common checkout price $p$---the total payment
at the register---and a per-unit tax $\tau$.  Fix $p$ and $\tau$.  Seller
remittance clears at
\begin{equation}
  D(p) = S\bigl(p - \tau - r(p)(p+h)\bigr),
\end{equation}
because both the return rate and per-return cost $p+h$ are evaluated at
checkout price $p$.  Buyer remittance clears at
\begin{equation}
  D(p) = S\bigl((p-\tau) - r(p-\tau)(p-\tau+h)\bigr),
\end{equation}
because the seller posts list price $p-\tau$, so returns and refund cost are
evaluated there.  The two conditions define the same equilibrium only if the
arguments of $S(\cdot)$ coincide.  Canceling $p-\tau$ yields
\begin{equation}
  r(p)(p+h) = r(p-\tau)(p-\tau+h).
\end{equation}

\begin{result}[Failure of tax neutrality with returns]
  Seller and buyer taxes define the same equilibrium if and only if
  $r(p)(p+h)=r(p-\tau)(p-\tau+h)$.\footnote{Neutrality requires the return
  wedge to be the same at checkout price $p$ and at posted list price
  $p-\tau$.  That holds for all taxes only if $r(p)(p+h)$ is constant in
  posted list price, which requires $r(p)=C/(p+h)$ for some constant $C$;
  without that functional form on $r$, the two clearing conditions need not
  define the same equilibrium.}
\end{result}

Tax neutrality breaks down because the refunded amount---and therefore the
keep-or-return decision---depends on which party remits the tax.  Under seller
remittance the tax-inclusive checkout price is refunded; under buyer remittance
only the pre-tax list price is.  Purchase is always based on checkout price
$p$, but returns are evaluated at $p$ under seller remittance and at $p-\tau$
under buyer remittance, so effective revenue generally differs and the two
clearing conditions are not equivalent.

%=============================================================================
\section{Pass-through}
%=============================================================================

Equilibrium satisfies $D(p)=S(\mu(p,h))$.  Write $Q\equiv D(p)$.  We
characterize pass-through into list price for three shocks: a uniform
one-dollar parallel shift in every firm's marginal production cost (indexed by
$c$), a one-dollar increase in handling cost $h$, and a unit tax under seller
and buyer remittance.  Production cost shifts market supply at a given
effective margin $\mu$, while handling lowers $\mu(p,h)$ at a fixed list
price $p$.

For production and handling, write the clearing condition as
\begin{equation}
  m(p,x)\equiv D(p)-S(\mu(p,x),x)=0,
  \qquad x\in\{c,h\}.
\end{equation}
Differentiating while keeping $m=0$ defines pass-through $\rho_x\equiv dp/dx$ by
\begin{equation}
  \rho_x=-\frac{m_x}{m_p},
  \label{eq:pass-through-general}
\end{equation}
and expanding $m=D-S$ gives
\begin{align}
  m_p &= D'(p)-S'(\mu)\,\mu_p,
  &
  m_x &= -S'(\mu)\,\mu_x-S_x,
\end{align}
hence
\begin{equation}
  \rho_x
  = \frac{-S'(\mu)\,\mu_x-S_x}{S'(\mu)\,\mu_p-D'(p)}.
  \label{eq:pass-through}
\end{equation}
The two shocks differ in which margin moves at fixed $p$ or fixed $\mu$,
\begin{align}
  \mu_c &= 0,
  &
  S_c &= -S'(\mu),
  &
  \mu_h &= -r(p),
  &
  S_h &= 0.
\end{align}
The production term is $S_c=-S'(\mu)$ because a one-dollar parallel rise in
every firm's marginal cost is equivalent, at fixed $\mu$, to evaluating market
supply at $\mu-1$.  Handling moves $\mu$ by only $r(p)$ per dollar relative to
that unit production shift, so its numerator in \eqref{eq:pass-through} is the
fraction $r(p)$ of the production numerator.

Turning to the unit tax, failure of tax neutrality requires separate checkout
prices $p^s$ and $p^b$ under seller and buyer remittance.  Write
$\rho^s_\tau\equiv dp^s/d\tau$ and $\rho^b_\tau\equiv dp^b/d\tau$.  At
$\tau=0$, \eqref{eq:pass-through} with $S_\tau=0$ gives
\begin{equation}
  \rho^s_\tau
  = \frac{-S'(\mu)\,\mu^s_\tau}{S'(\mu)\,\mu_p-D'(p^s)},
  \qquad
  \rho^b_\tau
  = \frac{-S'(\mu)\,\mu^b_\tau}{S'(\mu)\,\mu_p-D'(p^b)}.
\end{equation}
The regimes differ only through $\mu^s_\tau$ and $\mu^b_\tau$ at fixed
checkout price.  At $\tau=0$, $\mu^s_\tau=-1$.  Under buyer remittance,
$\tau$ enters effective revenue in the same way as $-p$, so
$\mu^b_\tau=-\mu_p$.  Seller remittance therefore has the same format as a
production-cost shock, while buyer remittance scales that format by $\mu_p$.
Collecting these cases, we have
\begin{result}[Pass-through into list price]
  Production-cost pass-through is
  \begin{equation}
    \rho_c
    = \frac{S'(\mu)}{S'(\mu)\,\mu_p-D'(p)}.
    \label{eq:rho-c}
  \end{equation}
  The remaining shocks connect to this object by
  \begin{align}
    \rho_h &= r(p)\,\rho_c,
    &
    \rho^s_\tau &= \rho_c,
    &
    \rho^b_\tau &= \mu_p\,\rho_c.
    \label{eq:rho-links}
  \end{align}
\end{result}
When $r\equiv 0$, one has $\mu_p=1$ and therefore
$\rho_c=\rho^s_\tau=\rho^b_\tau$, which equals the textbook competitive
pass-through.

\subsection{Elasticity representation}

The same pass-through formulas can be rewritten in elasticities, which makes
the role of $\mu_p$ transparent.  Write $\mu_p\equiv d\mu/dp$ for the response
of effective revenue to list price, and define demand and supply elasticities
with respect to list price as usual by
\[
\varepsilon_D(p)\equiv\frac{p D'(p)}{D(p)}<0,
\qquad
\varepsilon_S(p)\equiv\frac{dS(\mu(p))}{dp}\frac{p}{S(\mu(p))}
=S'(\mu)\,\mu_p\,\frac{p}{S(\mu)}.
\]
Here $\varepsilon_S(p)$ is the observed elasticity of market supply with
respect to the clearing price.  Under $S'(\mu)>0$, it has the sign of
$\mu_p$.

Without returns, supply is a function of list price, so $S'=dS/dp$.  A
one-dollar cost increase cuts supply by $S'$ units at the original price,
creating an excess-demand gap of that size.  Raising list price by one dollar
closes the gap from both sides.  Demand falls by $|D'|$ and supply recovers by
$S'$.  Closing a gap of size $S'$ therefore requires a price rise of
$S'/(S'+|D'|)$, so standard pass-through equals
$\varepsilon_S/(-\varepsilon_D+\varepsilon_S)$.

With returns, supply is a function of effective revenue, so
$S'(\mu)=dS/d\mu$ rather than $dS/dp$.  A one-dollar production-cost increase
is equivalent to lowering $\mu$ by one dollar, so at fixed list price supply
falls by $S'(\mu)$ units, again creating an excess-demand gap of that size.
Raising list price by one dollar closes the gap at rate
$S'(\mu)\,\mu_p-D'(p)$, so
\[
\rho_c=\frac{S'(\mu)}{S'(\mu)\,\mu_p-D'(p)}.
\]
The cost shock lowers $\mu$ by one dollar, while a dollar of list price raises
$\mu$ by $\mu_p$.  A higher $\mu_p$ means a one-dollar price increase restores
more effective revenue---and therefore more quantity supplied---than the
one-dollar cost shock removes, so a smaller list-price movement closes the
excess-demand gap.  Measuring supply against list price through
$\varepsilon_S$ puts the factor $1/\mu_p$ in the numerator,
\begin{equation}
  \rho_c
  =
  \frac{\varepsilon_S(p)/\mu_p}
       {-\varepsilon_D(p)+\varepsilon_S(p)}.
  \label{eq:rho_c_elasticity_final}
\end{equation}
Pass-through is therefore dampened relative to the no-return benchmark when
$\mu_p>1$ and amplified when $\mu_p<1$.  If returns are absent, $r\equiv 0$
forces $\mu_p=1$, and \eqref{eq:rho_c_elasticity_final} collapses to the
standard formula.

Whether $\mu_p$ exceeds one depends on how steeply the return rate falls with
price.  Differentiating $\mu=p-r(p)(p+h)$ gives
\[
\mu_p=1-r(p)-r'(p)(p+h).
\]
Write $\varepsilon_r(p)\equiv p\,r'(p)/r(p)$ for the elasticity of the return rate
in list price.  Then $\mu_p>1$ if and only if $\varepsilon_r(p)<-p/(p+h)$.
That is, selection must outweigh the behavioral rise in individual return
probabilities by enough that aggregate $r$ falls steeply in $p$.  When instead
the behavioral response dominates, $\varepsilon_r$ is less negative (or
positive) and $\mu_p<1$.  Comparing $\mu_p$ with the classical benchmark
$\mu_p=1$ then signs the bias for an analyst who ignores returns.
\begin{result}
  Relative to the correct returns-adjusted $\rho_c$, the classical no-return
  pass-through formula overestimates pass-through when
  $\varepsilon_r(p)<-p/(p+h)$ and underestimates it otherwise.
\end{result}
The practical value is that the bias can be signed without estimating the full
returns-adjusted model.  The return-rate elasticity and the ratio of handling
to list price are enough.  The economics behind the threshold is familiar from
the selection and behavioral forces already in $\varepsilon_r$.  When selection
dominates, a higher list price lowers the return rate, so the
refund-and-handling drag on effective revenue shrinks as price rises.
Effective revenue can then rise by more than one dollar per dollar of list
price, and a smaller price movement closes the excess-demand gap than the
classical formula predicts.  Costly handling strengthens that channel and
makes overestimation more likely.  Otherwise, effective revenue rises by less
than one dollar per dollar of list price, so list price must move by more than
the classical formula predicts to close the gap, and the classical no-return
formula underestimates pass-through.

%=============================================================================
\section{Incidence}
%=============================================================================

Pass-through into list price determines the welfare loss from each cost shifter
and how that loss is split between consumers and producers.  Throughout,
$Q\equiv D(p)$ and $r\equiv r(p)$ denote the equilibrium gross quantity and the
return rate.  From the consumer-side analysis, a small increase in any shifter
$x$ changes consumer surplus by
\begin{equation}
  \frac{dCS}{dx} = -Q\,(1-r)\,\rho_x,
  \label{eq:CSx}
\end{equation}
where $\rho_x\equiv dp/dx$ is the pass-through of the shock into the list
price.

Producer surplus is less immediate because a shock may shift the supply
schedule directly as well as indirectly through the effective marginal revenue
$\mu$.  Let $S(\mu,x)$ be the market supply curve, where the second argument
allows for a direct effect of $x$ on firms' costs, and write producer surplus
as $PS(\mu,x)=\int_0^{\mu} S(z,x)\,dz$.  Differentiating with respect to $x$,
\[
  \frac{dPS}{dx}
  = S(\mu,x)\,\frac{d\mu}{dx}
    + \int_0^{\mu} \frac{\partial S(z,x)}{\partial x}\,dz .
\]
Using the equilibrium condition $S(\mu,x)=Q$ and denoting
$\Sigma_x\equiv\int_0^\mu (\partial S(z,x)/\partial x)\,dz$, we obtain
\begin{equation}
  \frac{dPS}{dx} = Q\,\frac{d\mu}{dx} + \Sigma_x
  = Q\bigl(\mu_p\,\rho_x + \mu_x\bigr) + \Sigma_x,
  \label{eq:PSx}
\end{equation}
with $\mu_p\equiv\partial\mu/\partial p$ and
$\mu_x\equiv\partial\mu/\partial x$ evaluated at fixed list price.

The term $\Sigma_x$ captures the direct shift of the supply curve holding
$\mu$ constant.  A uniform, parallel increase in every firm's marginal
production cost by one dollar ($x=c$) moves the supply curve inward according
to $\partial S(\mu,c)/\partial c = -S'(\mu)$, so
$\Sigma_c = \int_0^\mu (-S'(z))\,dz = -S(\mu) = -Q$.  For a handling-cost
shock $h$ or a per-unit tax (whether seller- or buyer-remitted), the supply
schedule at a given $\mu$ is unchanged, hence $\Sigma_h = \Sigma_{\tau_s} =
\Sigma_{\tau_b} = 0$.  The partial derivatives $\mu_x$ are obtained directly
from the definition of effective marginal revenue: $\mu_c=0$,
$\mu_h=-r$, $\mu_{\tau_s}=-1$, and, for the buyer-remitted tax evaluated at
$\tau=0$, $\mu_{\tau_b}=r+(p+h)r'-1$ as derived in Section~5.

Substituting these values together with the pass-through expressions
$\rho_h=r\rho_c$ and $\rho^s_\tau=\rho_c$ into \eqref{eq:PSx} yields the
producer-surplus responses:
\begin{align}
  \frac{dPS}{dc}   &= Q\bigl(\mu_p\,\rho_c - 1\bigr), \label{eq:PS-c} \\
  \frac{dPS}{dh}   &= Q\bigl(\mu_p\,\rho_h - r\bigr), \label{eq:PS-h} \\
  \frac{dPS}{d\tau_s} &= Q\bigl(\mu_p\,\rho_c - 1\bigr), \label{eq:PS-taus} \\
  \frac{dPS}{d\tau_b} &= Q\bigl(\mu_p\,\rho^b_\tau + \mu_{\tau_b}\bigr). \label{eq:PS-taub}
\end{align}

The incidence ratio $I(x)\equiv|dCS/dx|\,/\,|dPS/dx|$ measures how the total
loss is divided.  For production cost, handling cost, and the seller-remitted
tax, the ratio simplifies to the same expression:
\begin{equation}
  I(c)=I(h)=I(\tau_s)=\frac{(1-r)\,\rho_c}{1-\mu_p\,\rho_c}.
  \label{eq:I-symmetric}
\end{equation}
For the buyer-remitted tax, the ratio is
\begin{equation}
  I(\tau_b)=\frac{(1-r)\,\rho^b_\tau}
                  {-\bigl(\mu_p\,\rho^b_\tau + \mu_{\tau_b}\bigr)},
  \qquad \mu_{\tau_b}=r+(p+h)r'-1.
  \label{eq:I-taub}
\end{equation}

Equation \eqref{eq:I-symmetric} generalises the standard competitive-market
formula $I=\rho/(1-\rho)$ (Weyl and Fabinger, 2013).  Consumer loss is scaled
by the keep probability $1-r$, and the producer-side denominator involves the
effective-revenue pass-through $\mu_p\rho_c$ rather than the list-price
pass-through alone.  Production cost, handling cost, and a seller-remitted tax
all generate the same incidence split because none of them changes the price
at which consumers decide whether to return the product; they shift
equilibrium exclusively through the effective marginal revenue $\mu$.  The
buyer-remitted tax, in contrast, alters the refund price to the pre-tax list
price, so the return rate and the per-return refund liability differ.  This
breaks the symmetry, giving an incidence ratio that departs from
\eqref{eq:I-symmetric}---a welfare counterpart to the tax-neutrality failure
established earlier.  When returns are absent ($r=0$, $\mu_p=1$,
$\rho^b_\tau=\rho_c$), all four ratios collapse to $\rho_c/(1-\rho_c)$,
restoring the familiar equivalence.

\end{document}
