\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\title{Tax Pass-Through and Incidence:\\
Monopoly (Weyl and Fabinger, 2013) and Returns}
\author{Draft notes}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Self-contained notes on pass-through and tax incidence, following
Weyl and Fabinger (2013) with full derivations under perfect competition
and monopoly (Sections~\ref{sec:pc}--\ref{sec:monopoly}), then extended
to a monopoly with product returns, where $r(p)$ is the aggregate return
rate at price $p$ with $r'(p)>0$ (Section~\ref{sec:returns}, full derivations).
\end{abstract}

%=============================================================================
\section{Shared notation}
\label{sec:notation}
%=============================================================================

\begin{itemize}
  \item $Q = D(p)$: demand at consumer price $p$, with $D' < 0$.
  \item $\varepsilon_D \equiv -p D'(p)/D(p) > 0$: price elasticity of demand.
  \item $t$ or $\tau$: per-unit tax or cost shock borne by the seller.
  \item $\rho \equiv dp/d\tau$: \textbf{pass-through} (consumer price rise per
    dollar of seller cost).
  \item $CS(p) = \int_p^{\bar p} D(x)\,dx$, $\pi$: consumer surplus and profit.
  \item $I \equiv |\Delta CS|/|\Delta \pi|$: \textbf{incidence ratio} (monopoly);
    under perfect competition $I = \rho/(1-\rho)$.
\end{itemize}

%=============================================================================
\section{Perfect competition: Weyl and Fabinger, Section II}
\label{sec:pc}
%=============================================================================

Consumers pay $p$; producers receive $p - \tau$ when a per-unit tax $\tau$ is
levied on producers.  Market clearing is
\begin{equation}
  D(p) = S(p - \tau).
  \label{eq:pc-clearing}
\end{equation}
Let $Q$ denote equilibrium quantity and $\rho \equiv dp/d\tau$.

%-----------------------------------------------------------------------------
\subsection{Principle 1: Physical incidence is neutral}
%-----------------------------------------------------------------------------

If the tax is collected from consumers instead, $p$ is the price consumers
pay and producers receive $p - \tau$.  Equation \eqref{eq:pc-clearing} is
unchanged, so $(p,\,Q)$ is unchanged.  Labeling alone differs.

If a consumer tax lowers nominal $p$ by $\Delta\tau$, the producer's receipt
$p - \tau$ falls by $1 - \rho$ when a producer tax of size $\Delta\tau$ would
have raised $p$ by $\rho$.

%-----------------------------------------------------------------------------
\subsection{Principles 2--3: Burden sharing and incidence}
%-----------------------------------------------------------------------------

Consumer surplus $CS(p) = \int_p^{\bar p} D(x)\,dx$ satisfies
$dCS = -D(p)\,dp$, so
\begin{equation}
  \frac{dCS}{d\tau} = -D(p)\,\rho.
  \label{eq:pc-dcs}
\end{equation}
Producer surplus is $PS(u)$ with $u \equiv p - \tau$ and
$PS(u) = \int_0^u S(x)\,dx$, so $PS'(u) = S(u) = Q$.  Since
$u = p - \tau$,
\[
  \frac{dPS}{d\tau}
  = PS'(u)\left(\frac{dp}{d\tau} - 1\right)
  = Q(\rho - 1)
  = -(1-\rho)\,Q.
  \label{eq:pc-dps}
\]
\emph{Principle 2:} the mechanical tax bill $Q\,d\tau$ splits with weights
$\rho$ on consumers and $1-\rho$ on producers.

\emph{Principle 3:}
\[
  I \equiv \frac{|dCS/d\tau|}{|dPS/d\tau|}
  = \frac{\rho Q}{(1-\rho)Q}
  = \frac{\rho}{1-\rho}.
\]

%-----------------------------------------------------------------------------
\subsection{Principle 4: Pass-through formula}
%-----------------------------------------------------------------------------

Differentiate \eqref{eq:pc-clearing} with respect to $\tau$:
\[
  D'(p)\,\frac{dp}{d\tau}
  = S'(p-\tau)\left(\frac{dp}{d\tau} - 1\right).
\]
Collect terms:
\[
  \bigl[S'(p-\tau) - D'(p)\bigr]\,\rho = S'(p-\tau)
  \quad\Rightarrow\quad
  \rho = \frac{S'(p-\tau)}{S'(p-\tau) - D'(p)}.
\]
Define elasticities at the equilibrium:
\[
  \varepsilon_D \equiv -\frac{p\,D'(p)}{D(p)} > 0,
  \qquad
  \varepsilon_S \equiv \frac{(p-\tau)\,S'(p-\tau)}{S(p-\tau)} > 0.
\]
Since $Q = D(p) = S(p-\tau)$, multiply numerator and denominator by
$(p-\tau)/Q$:
\[
  \rho = \frac{\varepsilon_S}{\varepsilon_S + \varepsilon_D}
  = \frac{1}{1 + \varepsilon_D/\varepsilon_S}.
\]
The inelastic side bears more of the tax.

%-----------------------------------------------------------------------------
\subsection{Principle 5: Local to global}
%-----------------------------------------------------------------------------

For a finite tax change $\tau_0 \to \tau_1$, integrate
\eqref{eq:pc-dcs}--\eqref{eq:pc-dps}:
\[
  \Delta CS = -\int_{\tau_0}^{\tau_1} \rho(\tau)\,Q(\tau)\,d\tau,
  \qquad
  \Delta PS = -\int_{\tau_0}^{\tau_1} (1-\rho(\tau))\,Q(\tau)\,d\tau.
\]
Define the quantity-weighted average pass-through
\[
  \bar\rho_{\tau_0}^{\tau_1}
  \equiv
  \frac{\int_{\tau_0}^{\tau_1} \rho(\tau)\,Q(\tau)\,d\tau}
       {\int_{\tau_0}^{\tau_1} Q(\tau)\,d\tau}.
\]
Then
\[
  I_{\tau_0}^{\tau_1}
  = \frac{\Delta CS}{\Delta PS}
  = \frac{\bar\rho_{\tau_0}^{\tau_1}}{1 - \bar\rho_{\tau_0}^{\tau_1}}.
\]
The local formula $I = \rho/(1-\rho)$ holds globally once $\rho$ is replaced
by $\bar\rho$.  If $\tau$ rises until $Q=0$, the ratio $CS/PS$ of total
surpluses equals $\bar\rho/(1-\bar\rho)$ with $\bar\rho$ averaged from
$\tau=0$ to the choking tax.

%=============================================================================
\section{Monopoly: Weyl and Fabinger, Section III}
\label{sec:monopoly}
%=============================================================================

The monopolist chooses quantity $Q$ facing inverse demand $p(Q)$ with
$p'(Q) < 0$.  Revenue is $p(Q)\,Q$, cost is $c(Q)$ with
$\mathrm{MC}(Q) \equiv c'(Q)$.  Define
\[
  \mathrm{MR}(Q) \equiv \frac{d}{dQ}\bigl[p(Q)\,Q\bigr] = p(Q) + p'(Q)\,Q,
\]
\[
  m(Q) \equiv p(Q) - \mathrm{MC}(Q),
  \qquad
  ms(Q) \equiv -p'(Q)\,Q \quad \text{(marginal consumer surplus)}.
\]
With a per-unit producer tax $\tau$, the FOC is
\begin{equation}
  \mathrm{MR}(Q) = \mathrm{MC}(Q) + \tau.
  \label{eq:mr-tax}
\end{equation}
We write $p$, $Q$, $m$, $ms$ evaluated at the optimum unless noted.

%-----------------------------------------------------------------------------
\subsection{Principle 1: Physical incidence is neutral}
%-----------------------------------------------------------------------------

\paragraph{Producer tax.}
Problem: $\max_Q \bigl[p(Q) - \mathrm{MC}(Q) - \tau\bigr]\,Q$.
FOC: $\mathrm{MR}(Q) = \mathrm{MC}(Q) + \tau$.

\paragraph{Consumer tax.}
Consumers pay $p$; the monopolist keeps $p - \tau$.  Problem:
$\max_Q \bigl[p(Q) - \tau - \mathrm{MC}(Q)\bigr]\,Q$.
The FOC is identical.

\paragraph{Conclusion.}
Only $p - \tau - \mathrm{MC}$ matters for $Q$.  Physical incidence does not
affect real allocation; $\rho = dp/d\tau$ is the same under either labeling.

%-----------------------------------------------------------------------------
\subsection{Principle 2: Total burden exceeds tax revenue}
%-----------------------------------------------------------------------------

\paragraph{Step 1: Consumer burden.}
Demand is still $Q = D(p)$, so as in \eqref{eq:pc-dcs},
\begin{equation}
  \frac{dCS}{d\tau} = -Q\,\rho.
  \label{eq:mono-dcs}
\end{equation}

\paragraph{Step 2: Producer burden (envelope theorem).}
Profit is $\pi(Q,\tau) = \bigl[p(Q) - \mathrm{MC}(Q) - \tau\bigr]\,Q$.
Holding $Q$ fixed,
\[
  \frac{\partial \pi}{\partial \tau} = -Q.
\]
At the optimum, $\partial\pi/\partial Q = 0$, so the total derivative
equals the partial derivative:
\begin{equation}
  \frac{d\pi}{d\tau} = -Q.
  \label{eq:mono-dpi}
\end{equation}
The monopolist bears the full mechanical tax $Q\,d\tau$ at the margin.

\paragraph{Step 3: Compare to tax revenue.}
Revenue raised is $Q\,d\tau$.  Total burden on participants:
\[
  |\Delta CS| + |\Delta\pi|
  = \rho Q\,d\tau + Q\,d\tau
  = (1+\rho)\,Q\,d\tau
  > Q\,d\tau.
\]
Under perfect competition \eqref{eq:pc-dps}, producers bore
$(1-\rho)Q\,d\tau$ and the total equaled revenue.  Under monopoly the
extra $\rho Q\,d\tau$ is \textbf{excess burden} on consumers from the
quantity distortion.

%-----------------------------------------------------------------------------
\subsection{Principle 3: Incidence equals pass-through}
%-----------------------------------------------------------------------------

From \eqref{eq:mono-dcs}--\eqref{eq:mono-dpi},
\[
  I \equiv \frac{|\Delta CS|}{|\Delta\pi|}
  = \frac{\rho Q}{Q}
  = \rho.
\]
The monopolist pays the tax dollar-for-dollar (at fixed $Q$); any consumer
loss beyond mechanical shifting is excess burden, and its ratio to producer
loss is $\rho$.

%-----------------------------------------------------------------------------
\subsection{Social incidence of competition (constant marginal cost)}
%-----------------------------------------------------------------------------

Assume $\mathrm{MC} \equiv c$.  A quantity $\tilde q$ is supplied
competitively at cost $c$.  Incumbent profit:
$\pi = \bigl[p(Q) - c\bigr](Q - \tilde q)$.

\paragraph{Step 1: Competition mimics a cost shock.}
Marginal revenue becomes $p(Q) + p'(Q)(Q - \tilde q)$.  Increasing
$\tilde q$ by $d\tilde q$ shifts the FOC exactly like raising $\tau$ by
$-p'(Q)\,d\tilde q$:
\[
  \frac{dQ}{d\tilde q} = p'(Q)\,\frac{dQ}{d\tau}
  = \frac{p'(Q)}{\mathrm{MR}'} = \rho.
\]

\paragraph{Step 2: Deadweight loss.}
$m(Q) = p(Q) - c$.  Social optimum $Q^{**}$ satisfies $m(Q^{**}) = 0$.
\[
  \mathrm{DWL}(Q) = \int_Q^{Q^{**}} m(q)\,dq
  \quad\Rightarrow\quad
  \frac{d\,\mathrm{DWL}}{dQ} = -m(Q).
\]
Hence
\[
  \frac{d\,\mathrm{DWL}}{d\tilde q}
  = \frac{d\,\mathrm{DWL}}{dQ}\,\frac{dQ}{d\tilde q}
  = -m\,\rho.
\]

\paragraph{Step 3: Profit loss.}
Envelope theorem at fixed $Q$: $d\pi/d\tilde q = -m$.

\paragraph{Step 4: Social incidence.}
\[
  \mathrm{SI}
  \equiv \frac{d\,\mathrm{DWL}/d\tilde q}{d\pi/d\tilde q}
  = \frac{-\rho m}{-m}
  = \rho.
\]
Competition erodes deadweight loss and profit at the same rate as consumer
incidence of a tax.

%-----------------------------------------------------------------------------
\subsection{Principle 4: Pass-through formula}
%-----------------------------------------------------------------------------

\paragraph{Step 1: Differentiate the FOC.}
From \eqref{eq:mr-tax}, $\mathrm{MR}(Q) = \mathrm{MC}(Q) + \tau$.
Differentiate with respect to $\tau$:
\[
  \mathrm{MR}'(Q)\,\frac{dQ}{d\tau}
  = \mathrm{MC}'(Q)\,\frac{dQ}{d\tau} + 1
  \quad\Rightarrow\quad
  \bigl[\mathrm{MR}' - \mathrm{MC}'\bigr]\,\frac{dQ}{d\tau} = 1.
\]
Hence
\[
  \frac{dQ}{d\tau} = \frac{1}{\mathrm{MR}' - \mathrm{MC}'}.
\]

\paragraph{Step 2: Pass-through in price.}
Consumer price is $p(Q)$, so
\begin{equation}
  \rho = \frac{dp}{d\tau}
  = p'(Q)\,\frac{dQ}{d\tau}
  = \frac{p'(Q)}{\mathrm{MR}' - \mathrm{MC}'}.
  \label{eq:rho-general}
\end{equation}

\paragraph{Step 3: Constant marginal cost ($\mathrm{MC}' = 0$).}
$\mathrm{MR} = p + p'Q$, so
\[
  \mathrm{MR}' = p' + p' + p''Q = 2p' + p''Q,
\]
and \eqref{eq:rho-general} becomes
\begin{equation}
  \rho = \frac{p'}{2p' + p''Q}
  = \frac{1}{2 + p''Q/p'}.
  \label{eq:rho-curvature}
\end{equation}

\paragraph{Step 4: Introduce $\varepsilon_{ms}$.}
Since $ms = -p'Q$,
\[
  ms' = \frac{d}{dQ}(-p'Q) = -p' - p''Q
  \quad\Rightarrow\quad
  \frac{p''Q}{p'} = -\frac{ms'}{p'} - 1.
\]
Define
\[
  \varepsilon_{ms} \equiv \frac{ms}{ms'\,Q}
  \quad\Rightarrow\quad
  \frac{1}{\varepsilon_{ms}} = \frac{ms'\,Q}{ms}
  = \frac{p''Q}{p'} + 1,
\]
so $p''Q/p' = 1/\varepsilon_{ms} - 1$.  Substitute into
\eqref{eq:rho-curvature}:
\[
  \rho = \frac{1}{2 + 1/\varepsilon_{ms} - 1}
  = \frac{1}{1 + 1/\varepsilon_{ms}}.
\]
This is the \emph{curvature} part; supply and demand elasticities enter next.

\paragraph{Step 5: Introduce $\varepsilon_D$ and $\varepsilon_S$.}
Write \eqref{eq:rho-general} as
$\rho = 1/\bigl(\mathrm{MR}'/p' - \mathrm{MC}'/p'\bigr)$.
Since $\mathrm{MR} = p - ms$ with $ms \equiv -p'Q$,
\[
  \mathrm{MR}' = p' - ms'
  \quad\Rightarrow\quad
  \frac{\mathrm{MR}'}{p'}
  = 1 - \frac{ms'}{p'}
  = 2 + \frac{p''Q}{p'}
  = 1 + \frac{1}{\varepsilon_{ms}},
\]
using $ms'/p' = -1 - p''Q/p'$ and $p''Q/p' = 1/\varepsilon_{ms} - 1$.
For the cost side, with $\varepsilon_S \equiv \mathrm{MC}/(\mathrm{MC}'\,Q)$,
\[
  \frac{\mathrm{MC}'}{p'}
  = \frac{1}{\varepsilon_S\,Q}\cdot\frac{\mathrm{MC}}{p'}.
\]
At $\tau = 0$, Lerner's rule $\mathrm{MC}/p = (\varepsilon_D-1)/\varepsilon_D$ and
$\varepsilon_D = -p/(Q p')$ imply
\[
  \frac{\mathrm{MC}}{p'}
  = \frac{\mathrm{MC}}{p}\cdot\frac{p}{p'}
  = -\frac{(\varepsilon_D - 1)\,Q}{p},
  \qquad
  \frac{\mathrm{MC}'}{p'} = -\frac{\varepsilon_D - 1}{\varepsilon_S}.
\]
Therefore
\[
  \frac{\mathrm{MR}'}{p'} - \frac{\mathrm{MC}'}{p'}
  = 1 + \frac{1}{\varepsilon_{ms}} + \frac{\varepsilon_D - 1}{\varepsilon_S},
\]
and
\begin{equation}
  \rho = \frac{1}{1 + \dfrac{\varepsilon_D - 1}{\varepsilon_S}
    + \dfrac{1}{\varepsilon_{ms}}}.
  \label{eq:monopoly-rho}
\end{equation}
When $\mathrm{MC}' = 0$ (constant marginal cost), $\varepsilon_S \to \infty$ and
\eqref{eq:monopoly-rho} reduces to $\rho = 1/(1 + 1/\varepsilon_{ms})$ from
Step~4.

\paragraph{Step 6: Interpretation.}
\begin{itemize}
  \item \textbf{vs.\ perfect competition:} $\varepsilon_D$ becomes
    $\varepsilon_D - 1$ because monopoly requires $\varepsilon_D > 1$.
  \item \textbf{Curvature:} $1/\varepsilon_{ms}$ comes from $p''$.
    Log-concave demand ($\phi'' < 0$ for $\phi = \log D$) implies
    $\rho > 1$ when $\mathrm{MC}$ is constant.
  \item \textbf{Linear inverse demand} $p(Q) = a - bQ$: $p'' = 0$,
    $\mathrm{MR}' = -2b$, $p' = -b$, so $\rho = 1/2$.
\end{itemize}

%-----------------------------------------------------------------------------
\subsection{Principle 5: Local to global}
%-----------------------------------------------------------------------------

Monopoly formulas \eqref{eq:mono-dcs}--\eqref{eq:mono-dpi} integrate as in
Section~\ref{sec:pc}:
\[
  \Delta CS = -\int \rho(\tau)\,Q(\tau)\,d\tau,
  \qquad
  \Delta\pi = -\int Q(\tau)\,d\tau.
\]
With $\bar\rho$ the quantity-weighted average,
$I = \bar\rho$ (not $\bar\rho/(1-\bar\rho)$).

For social incidence with competition $\tilde q \in [0, Q^{**}]$, average
$\rho$ with weight $m(\tilde q)$:
\[
  \tilde\rho \equiv
  \frac{\int_0^{Q^{**}} \rho(\tilde q)\,m(\tilde q)\,d\tilde q}
       {\int_0^{Q^{**}} m(\tilde q)\,d\tilde q}.
\]
Then $\mathrm{DWL}/\pi = \tilde\rho$ globally (both numerator and
denominator vanish at $\tilde q = Q^{**}$).

%-----------------------------------------------------------------------------
\subsection{Summary: perfect competition vs.\ monopoly}
%-----------------------------------------------------------------------------

\begin{center}
\begin{tabular}{lcc}
  & Perfect competition & Monopoly \\
  \hline
  $dCS/d\tau$ & $-\rho Q$ & $-\rho Q$ \\
  $d\pi/d\tau$ & $-(1-\rho)Q$ & $-Q$ \\
  $I$ & $\rho/(1-\rho)$ & $\rho$ \\
  $\rho$ & $1/(1+\varepsilon_D/\varepsilon_S)$
  & $1/(1+\frac{\varepsilon_D-1}{\varepsilon_S}+\frac{1}{\varepsilon_{ms}})$ \\
\end{tabular}
\end{center}

%=============================================================================
\section{Monopoly with product returns}
\label{sec:returns}
%=============================================================================

This section extends the monopoly incidence framework to product returns.
We keep $\rho \equiv dp/d\tau$ and $I \equiv |\Delta CS|/|\Delta\pi|$ from
Section~\ref{sec:monopoly}.

%-----------------------------------------------------------------------------
\subsection{Environment}
%-----------------------------------------------------------------------------

\paragraph{Demand.}
At price $p$, quantity purchased is $Q = D(p)$ with $D'(p) < 0$.

\paragraph{Aggregate return rate.}
When the seller sets price $p$ and sells $Q = D(p)$ units, a share
$r(p) \in (0,1)$ of those units is returned.  Thus $r(p)$ is the
\textbf{aggregate (conditional) return rate}: among all units sold at
price $p$, fraction $r(p)$ come back.  We assume
\begin{equation}
  r'(p) > 0:
  \label{eq:r-increasing}
\end{equation}
the aggregate return rate is higher at higher prices (e.g.\ worse
price--quality match, more try-and-return when the tag price is high).

\paragraph{Gross returns.}
Total returned units are
\[
  G(p) \equiv D(p)\,r(p).
\]
When $p$ rises, $D(p)$ falls but $r(p)$ rises.  So $G(p)$ can move either
way; $r'(p)>0$ is the statement that the \textbf{return rate} itself
increases with price, not that total returns $G(p)$ must rise.

\paragraph{Payoffs.}
The seller produces $Q = D(p)$ units at cost $c$ and pays handling $h \ge 0$
per returned unit.  If kept: profit $(p-c)$.  If returned: refund $p$, cost
$c$, handling $h$, so profit $(-c-h)$.

\paragraph{Step 1: Derive $\mu(p)$.}
Among units sold at price $p$, share $1-r(p)$ are kept and share $r(p)$
are returned:
\begin{align*}
  \mathbb{E}[\text{profit per unit sold}]
  &= (1-r(p))(p-c) + r(p)(-c-h) \\
  &= p - c - r(p)(p+h).
\end{align*}
Define the \textbf{effective margin}
\begin{equation}
  \mu(p) \equiv p - c - r(p)(p+h).
  \label{eq:margin}
\end{equation}

\paragraph{Step 2: Total profit.}
\begin{equation}
  \max_p \;\pi(p) = D(p)\,\mu(p).
  \label{eq:profit-returns}
\end{equation}
The seller produces $D(p)$ units; aggregate return rate $r(p)$ implies
$G(p)=D(p)r(p)$ units returned.

%-----------------------------------------------------------------------------
\subsection{Pricing and the Lerner condition with returns}
%-----------------------------------------------------------------------------

\paragraph{Step 1: Derivative of $\mu$.}
Differentiate \eqref{eq:margin}:
\begin{equation}
  \mu'(p) = 1 - \underbrace{r(p)}_{\text{level effect}}
  - \underbrace{r'(p)\,(p+h)}_{\text{slope effect}}.
  \label{eq:mu-prime}
\end{equation}
\begin{itemize}
  \item \textbf{Level effect $-r(p)$:} a share $r(p)$ of sales are refunded, so
    the seller keeps only $(1-r(p))$ of the list price on average.
  \item \textbf{Slope effect $-r'(p)(p+h)$:} because $r'(p)>0$, raising $p$
    increases the aggregate return rate and erodes margin on the margin
    (pun intended).  This term is absent in standard Lerner.
\end{itemize}
When $r'(p)=0$ (return rate fixed in $p$), $\mu'(p)=1-r(p)<1$: returns lower
the marginal effective margin but do not make it fall further as $p$ rises.

\paragraph{Step 2: First-order condition.}
From \eqref{eq:profit-returns},
\begin{equation}
  \pi'(p) = D'(p)\,\mu(p) + D(p)\,\mu'(p) = 0.
  \label{eq:foc-returns}
\end{equation}
Rearrange:
\begin{equation}
  -\frac{D'(p)}{D(p)} = \frac{\mu'(p)}{\mu(p)}.
  \label{eq:foc-ratio}
\end{equation}
The left-hand side is the elasticity of demand (up to sign); the right-hand
side is the elasticity of the \textbf{effective margin} $\mu$ with respect
to price.

\paragraph{Step 3: Demand elasticity.}
By definition $\varepsilon_D \equiv -p\,D'(p)/D(p) > 0$, so
\begin{equation}
  -\frac{D'(p)}{D(p)} = \frac{1}{\varepsilon_D\, p}.
  \label{eq:demand-elasticity}
\end{equation}

\paragraph{Step 4: Lerner condition (general).}
Substitute \eqref{eq:demand-elasticity} into \eqref{eq:foc-ratio}:
\[
  \frac{1}{\varepsilon_D\, p} = \frac{\mu'(p)}{\mu(p)}
  \quad\Rightarrow\quad
  \boxed{\;\mu(p) = \varepsilon_D\, p\,\mu'(p).\;}
  \label{eq:lerner-returns}
\]
This is the \textbf{Lerner condition with returns}: at the optimum, the
effective margin $\mu$ equals demand elasticity times price times the
marginal change in $\mu$.

\paragraph{Step 5: Effective markup rate.}
Divide \eqref{eq:lerner-returns} by $p>0$:
\begin{equation}
  \frac{\mu(p)}{p}
  = \varepsilon_D\,\mu'(p).
  \label{eq:lerner-markup-rate}
\end{equation}
Define the \textbf{effective markup rate}
\[
  \mathcal{L}(p) \equiv \frac{\mu(p)}{p}
  = \frac{p - c - r(p)(p+h)}{p}
  = 1 - \frac{c}{p} - \frac{r(p)(p+h)}{p}.
\]
Then \eqref{eq:lerner-markup-rate} reads
$\mathcal{L} = \varepsilon_D\,\mu'$.
Compare to standard monopoly without returns ($r\equiv 0$):
\[
  \frac{p-c}{p} = \frac{1}{\varepsilon_D}
  \quad\Longleftrightarrow\quad
  \mu = \varepsilon_D\, p\,\mu' \text{ with } \mu'=1.
\]
Returns replace the simple markup $(p-c)/p$ by $\mathcal{L}$ and replace
$\mu'=1$ by $\mu' = 1 - r - r'(p+h)$.

\paragraph{Step 6: Decomposed Lerner identity.}
Substitute \eqref{eq:margin} and \eqref{eq:mu-prime} into
\eqref{eq:lerner-returns}:
\begin{equation}
  p - c - r(p)(p+h)
  = \varepsilon_D\, p\,\bigl[1 - r(p) - r'(p)(p+h)\bigr].
  \label{eq:lerner-expanded}
\end{equation}
Expand the right-hand side:
\[
  \varepsilon_D\, p - \varepsilon_D\, p\, r(p)
  - \varepsilon_D\, p\, r'(p)(p+h).
\]
Collect price terms on one side:
\begin{equation}
  p\bigl[1 - \varepsilon_D + \varepsilon_D\, r(p) + \varepsilon_D\, r'(p)(p+h)\bigr]
  = c + r(p)(p+h).
  \label{eq:lerner-price}
\end{equation}
\emph{Reading \eqref{eq:lerner-price}.}
\begin{itemize}
  \item \textbf{No returns} ($r=r'=0$): $p(1-\varepsilon_D)=c$, i.e.\
    $p = c/(1-1/\varepsilon_D)$, the usual monopoly markup formula.
  \item \textbf{Returns, fixed rate} ($r'(p)=0$): the bracket on the left is
    $1 - \varepsilon_D(1-r)$, so a higher aggregate return rate $r$ shrinks
    the net ``markup factor'' on $p$.
  \item \textbf{$r'(p)>0$}: the term $\varepsilon_D\, r'(p)(p+h)$ adds a
    further penalty on $p$, reflecting that a higher price raises $r(p)$ and
    tightens the optimum.
\end{itemize}

\paragraph{Step 7: Accounting vs.\ economic margin.}
It is tempting to plug $(p-c)/p$ into Lerner's rule, but that is wrong with
returns.  The \textbf{accounting} margin rate is $(p-c)/p$, while the
\textbf{economic} margin that enters the FOC is
\[
  \mathcal{L} = \frac{\mu}{p}
  = \frac{p-c}{p} - \frac{r(p)(p+h)}{p}.
\]
Returns create a wedge $r(p)(p+h)/p$ between list-price markup and profit
per dollar of price.  Optimal pricing equates $\mathcal{L}$ to
$\varepsilon_D\,\mu'$, not $(p-c)/p$ to $1/\varepsilon_D$.

\paragraph{Step 8: Constant aggregate return rate $r(p)=\alpha$.}
Then $\mu = (1-\alpha)p - c - \alpha h$ and $\mu' = 1-\alpha$.  Equation
\eqref{eq:lerner-returns} becomes
\[
  (1-\alpha)p - c - \alpha h = \varepsilon_D\, p\,(1-\alpha).
\]
Solve for $p$:
\begin{equation}
  p
  = \frac{c + \alpha h}{(1-\alpha)(1-\varepsilon_D)}.
  \label{eq:p-star-constant-r}
\end{equation}
(At an interior optimum, $\varepsilon_D>1$, so $1-\varepsilon_D<0$ and
$p>0$.)  The effective markup rate is
\[
  \mathcal{L} = \frac{\mu}{p} = \frac{1-\alpha}{\varepsilon_D},
\]
lower than $1/\varepsilon_D$ when $\alpha>0$: returns compress economic
markup at a given elasticity.

\paragraph{Step 9: Price-sensitive rate $r(p)=\alpha+\gamma p$.}
Then
\[
  \mu(p) = p - c - (\alpha+\gamma p)(p+h),
  \qquad
  \mu'(p) = 1 - \alpha - \gamma(2p+h).
\]
The Lerner condition \eqref{eq:lerner-expanded} is
\[
  p - c - (\alpha+\gamma p)(p+h)
  = \varepsilon_D\, p\,\bigl[1 - \alpha - \gamma(2p+h)\bigr].
\]
The left side is quadratic in $p$; the right side is also quadratic because
of the $\gamma(2p+h)$ term in $\mu'$.  In general $p^*$ solves a cubic after
substituting $\varepsilon_D = -p D'(p)/D(p)$.  With linear demand
$D(p)=a-bp$, $\varepsilon_D = bp/(a-bp)$ and a closed-form $p^*$ exists but
is lengthy; the qualitative point is that $r'(p)=\gamma>0$ enters both the
level of $\mu$ and the marginal $\mu'$, shifting $p^*$ downward relative to
the constant-$r$ case.

\paragraph{Step 10: Summary of Lerner with returns.}
\begin{center}
\begin{tabular}{lcc}
  \hline
  & Standard monopoly & With returns \\
  \hline
  Profit per unit & $p-c$ & $\mu = p-c-r(p+h)$ \\
  Marginal $\mu$ & $1$ & $1-r-r'(p+h)$ \\
  FOC / Lerner & $(p-c)/p = 1/\varepsilon_D$
  & $\mu/p = \varepsilon_D\,\mu'$ \\
  Markup object & $(p-c)/p$ & $\mathcal{L}=\mu/p$ \\
  \hline
\end{tabular}
\end{center}
Returns preserve the \emph{structure} of Lerner (elasticity $\times$ marginal
margin) but replace price-cost markup with the effective margin $\mu$ and its
derivative $\mu'$, which encode both the level $r(p)$ and slope $r'(p)$ of
the aggregate return rate.

%-----------------------------------------------------------------------------
\subsection{Pass-through of production cost $\tau$}
%-----------------------------------------------------------------------------

A per-unit cost shock gives $\mu(p,\tau) = p - c - \tau - r(p)(p+h)$ and
$\partial\mu/\partial\tau = -1$.

\paragraph{Step 1: Implicit equation.}
$f(p,\tau) \equiv D'\mu + D\mu' = 0$.

\paragraph{Step 2: Partial derivatives.}
$\partial f/\partial\tau = -D'(p)$;
\[
  \frac{\partial f}{\partial p}
  = D''\mu + 2D'\mu' + D\mu'',
  \quad
  \mu'' = -2r' - r''(p+h).
\]

\paragraph{Step 3: Pass-through.}
\begin{equation}
  \rho_\tau \equiv \frac{dp}{d\tau}
  = \frac{D'(p)}{D''\mu + 2D'\mu' + D\mu''}.
  \label{eq:rho-tau}
\end{equation}
Since $D'<0$ and the denominator is negative at a maximum, $\rho_\tau > 0$.

\paragraph{Step 4: No returns.}
If $r \equiv 0$, then $\mu = p-c-\tau$, $\mu'=1$, $\mu''=0$, and
\eqref{eq:rho-tau} equals $D'/(D''(p-c-\tau)+2D')$.

\paragraph{Step 5: Role of $r'(p)>0$.}
Returns affect $\rho_\tau$ through $\mu$ (level), $\mu'<1$ (slope), and
$\mu''$ (curvature of $r$).  No closed form like \eqref{eq:monopoly-rho}
without specifying $D$ and $r$.

%-----------------------------------------------------------------------------
\subsection{Pass-through of handling cost $h$}
%-----------------------------------------------------------------------------

From \eqref{eq:margin}, $\partial\mu/\partial h = -r(p)$: handling cost
applies to the fraction $r(p)$ of sold units that are returned.  Same derivation as above with $s(p)=r(p)$:
\begin{equation}
  \rho_h \equiv \frac{dp}{dh}
  = \frac{D'(p)\,r(p)}{D''\mu + 2D'\mu' + D\mu''}.
  \label{eq:rho-h}
\end{equation}
Pass-through of $h$ is scaled by the aggregate return rate $r(p)$.

%-----------------------------------------------------------------------------
\subsection{Incidence}
%-----------------------------------------------------------------------------

\paragraph{Step 1: Producer (envelope).}
For $\partial\mu/\partial\theta = -s(p)$ with $s=1$ for $\tau$ and $s=r$ for
$h$:
\[
  \frac{d\pi}{d\theta} = -D(p)\,s(p).
  \label{eq:dpi-theta}
\]

\paragraph{Step 2: Consumers.}
\[
  \frac{dCS}{d\theta} = -D(p)\,\rho_\theta.
  \label{eq:dcs-theta}
\]

\paragraph{Step 3: Incidence ratio.}
\begin{equation}
  I_\theta = \frac{|dCS/d\theta|}{|d\pi/d\theta|}
  = \frac{\rho_\theta}{s(p)}.
  \label{eq:incidence-returns}
\end{equation}
Production shock: $I_\tau = \rho_\tau$.  Handling shock: $I_h = \rho_h/r(p)$.

\paragraph{Step 4: Total burden.}
$|dCS| + |d\pi| = (\rho_\theta + s(p))\,D(p)\,d\theta > s(p)\,D(p)\,d\theta$
when $\rho_\theta > 0$: excess burden on consumers as in Section~\ref{sec:monopoly}.

%-----------------------------------------------------------------------------
\subsection{Example A: constant $r(p)=\alpha$, linear demand}
%-----------------------------------------------------------------------------

Let $r(p)=\alpha \in (0,1)$ and $D(p)=a-bp$, $b>0$.

\paragraph{Margins.}
$\mu = (1-\alpha)p - c - \alpha h$, $\mu' = 1-\alpha$, $\mu''=0$.

\paragraph{Pass-through.}
\[
  \rho_\tau = \frac{-b}{-2b(1-\alpha)} = \frac{1}{2(1-\alpha)} > \frac{1}{2},
  \qquad
  \rho_h = \frac{-b\,\alpha}{-2b(1-\alpha)} = \frac{\alpha}{2(1-\alpha)}
  = \alpha\,\rho_\tau.
\]

\paragraph{Incidence.}
$I_\tau = \rho_\tau$ and $I_h = \rho_h/\alpha = \rho_\tau$.

%-----------------------------------------------------------------------------
\subsection{Example B: price-sensitive aggregate return rate $r(p)=\alpha+\gamma p$}
%-----------------------------------------------------------------------------

With $\gamma>0$, the aggregate return rate rises with $p$: $r'(p)=\gamma>0$.
Then
\[
  \mu(p) = p - c - (\alpha+\gamma p)(p+h),
\]
\[
  \mu'(p) = 1 - \alpha - \gamma(2p+h) < 1,
  \qquad
  \mu''(p) = -2\gamma.
\]
Substitute into \eqref{eq:rho-tau}--\eqref{eq:rho-h} at the optimum $p^*$
from \eqref{eq:foc-returns}.  Compared to Example~A, $\rho_\tau$ is lower
when $\gamma$ is large: raising $p$ both reduces $D(p)$ and raises the
aggregate return rate $r(p)$.

%-----------------------------------------------------------------------------
\subsection{Summary}
%-----------------------------------------------------------------------------

\begin{center}
\begin{tabular}{ll}
  \hline
  Object & Formula \\
  \hline
  Aggregate return rate at $p$ & $r(p)$, $r'(p)>0$ \\
  Gross returns & $G(p) = D(p)\,r(p)$ \\
  Margin & $\mu = p - c - r(p)(p+h)$ \\
  FOC / Lerner & $\mu = \varepsilon_D\, p\,\mu'$
    \eqref{eq:lerner-returns} \\
  $\rho_\tau$ & $D'/(D''\mu + 2D'\mu' + D\mu'')$ \\
  $\rho_h$ & $D'\,r/(D''\mu + 2D'\mu' + D\mu'')$ \\
  Incidence & $I_\theta = \rho_\theta/s$, $s=1$ or $s=r$ \\
  \hline
\end{tabular}
\end{center}

%=============================================================================
\section*{References}
\addcontentsline{toc}{section}{References}
\begin{thebibliography}{9}

\bibitem[Weyl and Fabinger(2013)]{weyl2013pass}
Weyl, E.~G. and M. Fabinger (2013).
\newblock Pass-through as an economic tool: Principles of incidence under
  imperfect competition.
\newblock \emph{Journal of Political Economy}, 121(3):528--583.

\end{thebibliography}

\end{document}
